% !TeX root = ../main.tex

\begin{acknowledgements}
	First of all, I am very grateful to Heidelberg Univerisity (HU) and Univeristy of Science and Technology of China (USTC) for giving me the chance to write the thesis in Germany, and China Scholarship Council for offering me finicial support in this period of time.

	Then I really thank Professor Andreas Dreuw.
	When I first came to this group, he arranged a research project that was very suitable for me according to my research interests, which is ADC.
	This is very helpful for me to understand many fields related to it.
	I am also very grateful to Sebastian and Dirk.
	Sebastian gave me a lot of help in the implementation of the program, giving me a detailed explanation of how qchem works, and answering many of my specific questions about coding.
	With a lot of knowledge, Dirk also gave me many good advice when neither Sebistian nor I can solve some problems.

	I also need to thank many people at USTC.
	Professor Jun Jiang, who serves as my class monitor, supports me a lot when I was still a freshman and had little idea of what I should do and can do in the future.
	He is also the one that took me to the way of theoretical chemistry, which is proved to be the best decision I made in my undergraduate life.
	By working in his group on an application of embedding theory on graphene, I had chance to learn much in theoretical chemistry, which keeps me excited in science and research.
	Yukai, who is currently a PhD student at Princeton Univeristy and one year senior than me at USTC, gave me a lot of constructive suggestions on study and application for U.S. universities, and is also a person that I respect very much for his taste and understanding in physics.

	I also have to mention my best friends Linhao and Shubo.
	Linhao discussed a lot of academic problems with me very frequently and thus helped me a lot on the way to science.
	Shubo shared a lot of his thoughts on study and life with me and helped me be awared of the meaning of my life.
	Both of them helped me a lot to relieve depression when I had difficulties for several times, especially on TOEFL exam and application for U.S. universities.
	I also thank my friends Zheru, Xuchen for a lot of meaningful chattings and discussions with me, Sen, Zihan, Daozheng, Wanying for having worked with me on scientific projects and making great contributions.

	Finally and most importanly, I need to thank my father and mother, who love and support me without any reservation from my birth.
\end{acknowledgements}
